---
output: github_document
bibliography: grateful-refs.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# elicitr

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
<!-- badges: end -->

Elicitr is an R package, used to aggregate elicitation data from googlesheets. 
This package follows two formal elicitation methods:

* Continuous method (adapted from delphi-protocol for expert elicitation)
* Categorical method (adapted from Vernet et al. (2020, Biological invasion))

## Installation

You can install the development version of elicitr from [GitHub](https://github.com/) with:

``` r
#install.packages("pak")
pak::pak("sgvignali/elicitr")
```

## Example

This simulated dataset can be used to unterstand how {elicitr} works.

Three types of variables are included here:

* "percentage" i.e. \[0,1\]
* "positive numbers" i.e. \[0,inf)\]
* "real numbers" i.e. \[(-inf,inf)\]

And three types of elicitations:

* One point estimate (var1_best)\
* Three point estimate (var2_min, var2_max, var2_best)\
* Four point estimate (var3_min, var3_max, var3_best, var3_conf)\

```{r elicitr, echo = FALSE}
set.seed(25)
n <- 6
elicit <- data.frame(id = 1:n,
                     # 1 point estimate with percentages
                     var1_best = runif(n,
                                       min = 0.7,
                                       max = 1),
                     # 3 points estimate with positive numbers
                     var2_best = truncnorm::rtruncnorm(n,
                                                       a = 0,
                                                       b = 100,
                                                       mean = 50,
                                                       sd = 10),
                     # 4 points estimate with natural numbers
                     var3_best = truncnorm::rtruncnorm(n,
                                                       a = -20,
                                                       b = 20,
                                                       mean = 0,
                                                       sd = 10)) |>
  dplyr::mutate(var2_min = var2_best - runif(n,
                                             min = 1,
                                             max = 10),
                var2_max = var2_best + runif(n,
                                             min = 1,
                                             max = 10),
                var3_min = var3_best - truncnorm::rtruncnorm(n,
                                                             a = -10,
                                                             b = 10,
                                                             mean = 0,
                                                             sd = 5),
                var3_max = var3_best + truncnorm::rtruncnorm(n,
                                                             a = -10,
                                                             b = 10,
                                                             mean = 0,
                                                             sd = 5),
                var3_conf = sample(50:100,
                                   size = n,
                                   replace = TRUE)) |>
  dplyr::select(id,
                var1_best,
                var2_min,
                var2_max,
                var2_best,
                var3_min,
                var3_max,
                var3_best,
                var3_conf) |>
  dplyr::mutate(var1_best = round(var1_best, digits = 1),
                dplyr::across(var2_min:var3_best, round))
print(elicit)
```


This is a basic example which shows you how to use {elicitr} to analyse this data:

```{r example}
library(elicitr)
#elicitate(elicit)
## fake function elicitate, replace with actual function
```

```{r pressure, echo = FALSE}
library(ggplot2)

ggplot2::ggplot(elicit) +
  ggplot2::geom_point(aes(x = id,
                             y = var2_best,
                             color = factor(id))) +
  ggplot2::geom_errorbar(aes(x = id,
                             y = var2_best,
                             ymin = var2_min,
                             ymax = var2_max,
                             color = factor(id),
                             width = 0.05)) +
  ggplot2::xlab("id") +
  ggplot2::ylab("estimates") +
  ggplot2::theme_bw()
```

```{r}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```
